```{r}
library(tidyverse)
library(tidymodels)
library(GGally)
library(lmtest)

```

```{r}
air = airquality
```

## **Task 1: Describe this dataset. How many variables and observations are there? Is there any missing data? Which variable is likely to be the response (Y) variable**  
#There are 153 observations and 6 variables.  There are missing data points for Ozone and Solar.R throughout. Ozone is likely to be the response Y variable.  

## **Task 2: Remove columns with missing data**  

```{r}
air2 = drop_na(air)
```

## **How many rows and columns remain in this new (air2) data frame?**  
# There are 111 rows and 6 columns.  

## **Task 3**  
Use ggpairs to develop a visualization of and and to calculate correlation for the combination of variables in this dataset**  

```{r}
ggpairs(air2)
```

Use the “ggcorr” function to develop a correlation matrix for the variables.

```{r}
ggcorr(air,label=TRUE)
```
## **Which variable is most strongly correlated with the “Ozone” variable?**  
Temp at 0.7  
## **Which variable is least strongly correlated with the “Ozone” variable?**  
Day at 0.  

## **Task 4: Plot Temp vs Ozone  
```{r}
ggplot(air2,aes(x=Temp, y=Ozone)) + geom_point()
```
## **Describe the relationship between “Temp” and “Ozone**  
There is a positive coorelation between temp and ozone.  As temperature (x) increases, ozone (y) increases. 

## ** Task 5: Use tidymodels to create a linear regression model using “Temp” to predict “Ozone**   
```{r}
modl = lm(Ozone ~ Temp, air2)
summary(modl)
```
## ** a) Discuss the quality of this model**  
R squared is 0.488  and adjusted in about the same. Nothing to note without comparing to other predictors.  
Slope is 2.4391 (positive regression line).  
pvalue is <2e-16 so it's way smaller than 0.05 so the predictor variable is significant.  

Build Tidymodel recipe  
```{r}
modl_recipe = recipe(Ozone ~ Temp, air2)
modl_recipe
```

```{r}
lm_model =
  linear_reg() %>%
  set_engine("lm")
```

```{r}
lm_workflow =
  workflow() %>%
  add_model(lm_model) %>%
  add_recipe(modl_recipe)
```

```{r}
lm_fit = fit(lm_workflow,air2)
```

## ** b) Use the code below to generate 95% confidence intervals for the coefficients**  
```{r}
confint(lm_fit$fit$fit$fit)
```
## ***. In what range does the slope coefficient likely fall?**   
1.96 and 2.91  

## **Task 6: Redo task 4 to include the regression line**  
```{r}
ggplot(air2,aes(x=Temp, y=Ozone)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

## **Task 7: Develop a prediction for “Ozone” when “Temp” is 80**  

There are two ways to do this: Manually write the formula below (probably shouldn't do that) and the below method.  
```{r}
testdata = data.frame(Temp = c(80))
predict(lm_fit, new_data = testdata)
```
Prediction for Ozone when temp is 80 is 47.48272  

## ** Task 8: Perform appropriate model diagnostics to verify whether or not the model appears to meet the four linear regression model assumptions**  

#1: The predictor and response variable have a linear relationship  
Based on the plotted data, the predictor and response variable have a linear relationship.  

#2: Model errors (residuals) are independent (recall that a residual is the difference between a predicted value and the actual value)  
```{r}
dwtest(lm_fit$fit$fit$fit)
```
Using the DW test, the p.value is much higher than 0.05, determining that the variables are independent.

#3 Model residuals exhibit constant variance   
```{r}
air3 = air2 %>% mutate(resid1 = lm_fit$fit$fit$fit$residuals)
ggplot(air3,aes(x=Temp, y=resid1)) + geom_point()
```
There's a slight pattern in the residuals - a slight U shape.

#4 Model residuals are Normally-distributed  
```{r}
ggplot(air3, aes(x=resid1)) + geom_histogram()
```
There is a dip in the middle of the histogram but the data is distributed somewhat normally.



