---
output:
  word_document: default
  html_document: default
---
## **Module 4: Random Forests**  
**Alex Deese**  

```{r}
library(tidyverse)
library(tidymodels)
library(caret)
library(gridExtra)
library(vip)
library(ranger)
library(skimr)
```

```{r}
drug = read_csv("drug_data-1.csv")
```

```{r}
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
#str(drug)
```

```{r}
drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"
```

```{r}
drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44",
"45_54", "55_64", "65_"))) %>%
mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
mutate(Education = factor(Education, labels =
c("Under16", "At16", "At17", "At18", "SomeCollege",
"ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
mutate(Country = factor(Country,
labels = c("USA", "NewZealand", "Other", "Australia",
"Ireland","Canada","UK"))) %>%
mutate(Ethnicity = factor(Ethnicity,
labels = c("Black", "Asian", "White", "White/Black", "Other",
"White/Asian", "Black/Asian"))) %>%
mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
select(-ID)
```

```{r}
#str(drug_clean)
```

```{r}
drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
names(drug_clean)
```

## **Task 1: Check for missing data in our "drug_clean" dataframe.  Implement a reasonable strategy to remove any missingness.**  

There is no missing data in our dataframe.  
```{r}
skim(drug_clean)
```

## **Task 2: Split the dataset into training and testing sets.**  
```{r}
set.seed(1234) 
drug_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine)
train = training(drug_split)
test = testing(drug_split)
```

## **Task 3: Create appropriate visualizations to examine the relationship between each variable and "Nicotine".  Use grid.arrange to organize these variables.  Comment on the relationship between each variable and "Nicotine".**  
**Age:** Age does appear to impact nicotine use.  The younger a person is, the more likely to use nicotine.  

**Gender:** Females are slightly more likely to use nicotine than males, but not by much.  

**Education:**  Education does appear to impact nicotine use.  In general, the lower the education level, the more likely a person is to use nicotine.  

**Country:** Nicotine use does vary by country.  For example, nicotine use is much less likely in New Zealand, and much more likely in Ireland.  

**Ethnicity:** Ethnicity does impact nicotine use.  For example, nicotine use is more likely for White/Asian ethnicities and less likely for Black ethnicities.  

**Nscore:** Nscore (Neuroticism) does not appear to impact nicotine use.  

**Escore:** Escore (extraversion) does not appear to impact nicotine use.  

**Oscore:** People with a higher Oscore are slightly more likely to use nicotine, though the difference between lower Oscores is not large.  This makes sense because Oscore represents a person's openness to new experiences.  

**Ascore:** People with a higher Ascore (agreeableness) are slightly less likely to use nicotine, though the difference between lower Ascores is not large.  

**Cscore:** People with a higher Cscore (conscientiousness) are slightly less likely to use nicotine, though the difference between lower Cscores is not large.  

**Impulsive:** People with a higher impulsiveness score are slightly more likely to use nicotine, though the difference between lower impulsiveness is not large (but larger than the differences for Nscore, Escore, Ascore, and Cscore).  

**SS:** People with a higher SS (sensation seeing) score are more likely to use nicotine.  

```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90))
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill")+ theme(axis.text.x = element_text(angle = 90))
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill")+ theme(axis.text.x = element_text(angle = 90))
grid.arrange(p1,p2,p3,p4)
```

```{r}
p5 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")+ theme(axis.text.x = element_text(angle = 90))
p6 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p7 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p8 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
grid.arrange(p5,p6,p7,p8)
```

```{r}
p9 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
p10 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p11 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
p12 = ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()
grid.arrange(p9,p10,p11,p12)
```

## **Task 4: Create a random forest model on the training set to predict Nicotine using all of the variables in the dataset.**  

```{r K folds}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)
```

```{r random forest}
drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>%
  set_engine("ranger", importance = "permutation") %>%
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

rf_grid = grid_regular(
  mtry(range = c(2, 8)),
  min_n(range = c(5, 20)),
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid
)
```

```{r Visualize forest}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```
```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")
```

## **Task 5: Use the best mtry and min_n values from Task 4 to finalize the workflow and fit the model to the training set. Example the variable importance. What variables are most important to this model?**  

The best values from Task 4 are min_n of 16 and and mtry of 6.  

In this model, SS score is by far the most important variable to determine nicotine use.  Next in importance is Oscore (openness), Country_UK, Age_X45-54, and Cscore (conscientiousness). The remaining variables have a lower importance.  
 
```{r find best_rf}
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```

```{r fit to training set}
final_rf_fit = fit(final_rf, train)
```

```{r variable importance}
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

## **Task 7: How does the model perform on the training and testing sets?**  

The model has an accuracy of 0.9165 on the training set, compared to a naive accuracy of 0.6707.  This model is significantly more accurate than the naive model.  

The model is only 0.6966 accurate on the test set, compared to a naive accuracy of 0.6702.  Though the accuracy is higher than the naive value, it is not very good compared to the accuracy of 0.9165 on the train data set.  This may be evidence of overfitting on the training data set.  

```{r train pred}
trainpredrf = predict(final_rf_fit, train)
head(trainpredrf)
```

```{r train confusion matrtix}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```

```{r train test and confusion matrix}
testpredrf = predict(final_rf_fit, test)
head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

## **Task 7:  Comment on how this model might be used in the “real-world.” Would you recommend this model for real-world use? What if any concerns would you have about using the model?**  

This model might be used in the real-world to determine at-risk people for nicotine use. For example, to identify groups that may be more prone to using nicotine based on their age, demographic, personality, etc.  After identifying these groups, the results could be used to put programs in place to help educate and prevent the use of nicotine by these people.  On a darker note, these predictions could be used by tobacco companies to target potential customers (those that are more likely to use nicotine).  The results could also be used when determining health care costs (individuals that fall within a group more prone to nicotine use would have a higher cost for healthcare).  

I would have concerns using this model in real-world situations, because the accuracy of the model on the test data was not very high and was significantly different than the accuracy of the model on the training data.  
